A Machine Learning Approach to MLB Catcher Framing

### 2021-12-14 Paper.Rmd
Started working on logistic regression/ridge regression/LASSO models

### Logistic Regression

### Ridge Regression

### LASSO

>>>>

## Logistic Regression

```{r}
set.seed(445)
load("~/DSCI 445 proj/project-7/Data/usable_data.Rda")
data_no_catchers <- usable_data[, -18]

train <- sample(1:nrow(data_no_catchers), nrow(data_no_catchers)/2)
test <- (-train)

log_reg_fit <- glm(strike ~ ., data=data_no_catchers[train, ], family='binomial')
summary(log_reg_fit)

pred <- predict.glm(log_reg_fit, type='response', newdata = data_no_catchers[test, ])
pred[pred >= 0.5] <- 'strike'
pred[pred < 0.5] <- 'ball'

conf_mx <- table(pred=pred, true=data_no_catchers[test, ]$strike)
error <- (conf_mx[2] + conf_mx[3])/length(pred)

conf_mx
1-error
```

We started with a logistic regression classifier using the selected variables as a preliminary model for predicting strikes. The model did not perform very impressively, with a prediction accuracy of 0.67. The summary for the logistic regression model gives which predictors are statistically significant, which included: 
- Pitch type, release speed, z release position
- Whether the batter is a lefty/righty
- Count
- Where the pitch landed
- Most of the innings
- Batter height
- Whether it was a home game

## Ridge Regression and LASSO

#Ridge Regression

```{r}
set.seed(445)
library(glmnet)

x <- model.matrix(strike ~ ., data = data_no_catchers)[, -1]
y <- data_no_catchers$strike

lambda <- seq(0.001, 10, length=100)

#Create training and test data set
train <- sample(1:nrow(x), nrow(x)/2)
test <- (-train)
ridge_reg_fit <-  glmnet(x[train,], y[train], alpha = 0, lambda = lambda, family='binomial')

#Choose lambda
cv_ridge <- cv.glmnet(x[train, ], y[train], alpha=0, family='binomial')
lambda_choice <- cv_ridge$lambda.min

#Get predictions
ridge_predict <- predict(ridge_reg_fit, s = lambda_choice, newx = x[test, ], type = 'class')

#Confusion matrix
ridge_conf_mx <- table(pred=ridge_predict, true=y[test])
ridge_conf_mx
error <- (ridge_conf_mx[2] + ridge_conf_mx[3])/length(ridge_predict)
1-error
```
The ridge regression model performed modestly with a prediction accuracy of 0.672. This is hardly any different from the logistic regression model, so it also is not the best choice of model to be used to rank the catchers.

#LASSO
